Parsing Postfix log files

### 2009-02-16 doc/logparser-architecture.tex
Improved the framework section in the architecture chapter; it still needs work, but I'm moving on to actions.

To avoid cluttering the explanation of the parser architecture with details
of the difficulties encountered while parsing Postfix log files, the
description of the parser architecture has been separated from the
description of \parsername{}, the implementation developed for this
project.  This chapter presents the architecture that has been designed and
developed for this project, beginning with the overall architecture and
design, followed by the three components of the architecture: Framework,
Actions, and finally Rules.  The discussion in this chapter centers on the
theoretical, XXX EXTEND WHEN THE CHAPTER IS FINISHED\@.

The practical difficulties of writing a parser for Postfix log files are
covered in detail in \sectionref{Postfix Parser Implementation}.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                \\
        \tablemiddleline{}%
        Networks        & Algorithm     & Determines the sequence 
                                        of transitions              \\
                        &               & or actions that 
                                        constitutes a valid  input. \\
        Transitions     & Actions       & Save data and impose
                                        conditions the              \\
                        &               & input must meet to be
                                        accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for 
                                        classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

XXX PULL CONTENT FROM THE PAPER FOR HERE\@; IT IS BETTER WRITTEN\@.

XXX EXTEND FRAMEWORK SECTION IF NECESSARY\@.

@@

The intermingling of log entries from different mails immediately rules out
the possibility of handling each mail in isolation; the parser must be able
to handle multiple mails in parallel, each potentially at a different stage
in its journey, without any interference between mails --- except in the
minority of cases where intra-mail interference is required.  The best way
to implement this is to maintain state information for every unfinished
mail and manipulate the appropriate mail correctly for each log line
encountered.

This functionality is provided by the framework, which both drives parsing
overall and provides services to the actions it invokes.  The framework is
responsible for loading rules from the database and sanity checking them,
reading log files, matching each rule against the current log line,
invoking the correct action, maintaining state tables, loading and saving
state, displaying a progress indicator, and miscellaneous other tasks.
Actions use services provided by the framework, including storing and
retrieving data in the state tables, extracting and saving data captured by
rules, storage of global data, debugging functions, preparing a mail for
entry in the database, and entering mails in the database.

There is a similarity between this design and the event-driven programming
paradigm commonly used in GUI programs, where one part of the program
responds to events (mouse clicks in a GUI program, log lines being matched
in the parser) and invokes the correct action.

The framework's high level overview could be expressed as (indentation
denoting flow of control):

\begin{verbatim}
for each line in the input files:
    for each rule defined by the user:
        if this rule matches the input line:
            perform the action specified by the rule
            skip the remaining rules
            process the next input line
    warn the user that the input line was not parsed
\end{verbatim}

>>>>

To avoid cluttering the explanation of the parser architecture with details
of the difficulties encountered while parsing Postfix log files, the
description of the parser architecture has been separated from the
description of \parsername{}, the implementation developed for this
project.  This chapter presents the architecture that has been designed and
developed for this project, beginning with the overall architecture and
design, followed by the three components of the architecture: Framework,
Actions, and finally Rules.  The discussion in this chapter centers on the
theoretical, XXX EXTEND WHEN THE CHAPTER IS FINISHED\@.
The practical difficulties of writing a parser for Postfix log files are
covered in detail in \sectionref{Postfix Parser Implementation}.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                  \\
        \tablemiddleline{}%
        Networks        & Overall       & Determines the sequence 
                                          of transitions              \\
                        & Algorithm     & or actions that 
                                          constitutes a valid input.  \\
        Transitions     & Actions       & Assembles data and
                                          imposes conditions          \\
                        &               & the input must meet to be
                                          accepted as                 \\
                        &               & valid.                      \\
        Abbreviations   & Rules         & Responsible for 
                                          classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

XXX EXTEND FRAMEWORK SECTION\@: EXPAND ON ALL OF THE POINTS IN THE LIST IN
THE NEXT PARAGRAPH\@.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the programmers writing actions to concentrate
on writing productive code.  It links actions and rules, allowing either to
be improved independently of the other.  It provides shared storage to pass
data between actions, loads and saves state, loads and validates rules,
manages parsing, invokes actions, tracks how often each rule matches to
optimise rule ordering (\sectionref{rule ordering for efficiency}), stores
results of parsing, and miscellaneous other tasks.
Most parsers will require the same basic functionality from the framework,
plus some specialised support functions.  The framework is the core of the
architecture and is deliberately quite simple: the rules deal with the
variation in inputs, and the actions deal with the intricacies and
complications encountered when parsing.

The function that finds the rule matching the input and invokes the
requested action can be expressed in pseudo-code (indentation denoting flow
of control) as:

% DO NOT REFORMAT!

\begin{verbatim}
for each input:
    for each rule defined by the user: 
        if this rule matches the input:
            perform the action specified by the rule
            skip the remaining rules
            process the next input
    warn the user that the input was not parsed
\end{verbatim}

### 2009-02-16 doc/logparser-implementation.tex
Improved the framework section in the architecture chapter; it still needs work, but I'm moving on to actions.

\chapter{Postfix Parser Implementation}
\label{Postfix Parser Implementation}

@@

@@

The actions do the work of reconstructing the journey a mail takes through
Postfix.  Details of the actions available in the Postfix parser can be
found in \sectionref{actions in detail in implementation}, and
\sectionref{adding new actions in implementation} describes the process of
adding new actions.

>>>>

\chapter{Postfix Parser Implementation}

\label{Postfix Parser Implementation}

@@

The intermingling of log entries from different mails immediately rules out
the possibility of handling each mail in isolation; the parser must be able
to handle multiple mails in parallel, each potentially at a different stage
in its journey, without any interference between mails --- except in the
minority of cases where intra-mail interference is required.  The best way
to implement this is to maintain state information for every unfinished
mail and manipulate the appropriate mail correctly for each log line
encountered.

There is a similarity between this design and the event-driven programming
paradigm commonly used in GUI programs, where one part of the program
responds to events (mouse clicks in a GUI program, log lines being matched
in the parser) and invokes the correct action.

@@

The actions reconstruct the journey a mail takes through Postfix.  Details
of the actions available in the Postfix parser can be found in
\sectionref{actions in detail in implementation}, and \sectionref{adding
new actions in implementation} describes the process of adding new actions.
